---
title: "Analytics Edge Hwk3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Selected problems from Homework 3

### Loans dataset 

Loading dataset and installing packages used in balance of document
```{r}
loans = read.csv('loans.csv')
head(loans)
library(magrittr)
library(dplyr)
library(caTools)
library(zoo)
library(ROCR)
library(mice)
```

P1.4 (imputation)
```{r}
set.seed(144)
vars.for.imputation = setdiff(names(loans), "not.fully.paid")
imputed = complete(mice(loans[vars.for.imputation]))
loans[vars.for.imputation] = imputed
```

P1.5 (logistic regr)
```{r}
set.seed(144)
split = sample.split(loans$not.fully.paid,SplitRatio = 0.7)
train  = subset(loans, split)
test = subset(loans, !split)
loansMod1 = glm(not.fully.paid ~ ., data=loans, family='binomial')
summary(loansMod1)
```

P2.3. *Note: my answer for the first question does not  precisely match the official answer*
```{r}
predict.risk = predict(loansMod1, newdata = test, type = 'response')
table(test$not.fully.paid, predict.risk>=0.5)
1-mean(train$not.fully.paid)
```

P2.4
```{r}
ROCRpred = prediction(predict.risk, test$not.fully.paid)
as.numeric(performance(ROCRpred,'auc')@y.values)
```

P3.1
```{r}
mod_int.rate = glm(not.fully.paid~int.rate, data = train, family = 'binomial')
summary(mod_int.rate)
```

P3.2
```{r}
predictions_bivar = predict(mod_int.rate, type = 'response', newdata = test)
max(predictions_bivar)
table(test$not.fully.paid, predictions_bivar>=0.5)
```

P3.3
```{r}
ROCR_bivar_pred = prediction(predictions_bivar, test$not.fully.paid)
as.numeric(performance(ROCR_bivar_pred, 'auc')@y.values)
```

P5.1
```{r}
test$profit = exp(test$int.rate*3) - 1
test$profit[test$not.fully.paid == 1] = -1
max(test$profit*10)
```

P6.1
```{r}
test$predicted.risk = predict.risk
hiInterest=subset(test, int.rate>=0.15)
mean(hiInterest$profit)
mean(hiInterest$not.fully.paid)
```

P6.2 *Note: My answers don't match the official answers to the problem*
```{r}
cutoff = sort(hiInterest$predicted.risk, decreasing = F)[100]
selectedLoans = subset(hiInterest, predicted.risk<=cutoff)
(selectedLoans$profit)%>%sum
(selectedLoans$not.fully.paid)%>%sum
```

### Baseball Dataset

P1.4
```{r}
baseball = read.csv('baseball.csv')
madePlayoffs = subset(baseball, Playoffs==1)
nrow(madePlayoffs)
```

P1.5
```{r}
tapply(madePlayoffs$Team,madePlayoffs$Year,length)
```
P2.1/2.2/2.3/2.4 (**my answer to P2.4 doesn't match the official one**)
```{r}
playOffTbl=table(madePlayoffs$Year)
names(playOffTbl)
playOffTbl[c("1990","2001")]
baseball$NumCompetitors = playOffTbl[as.character(baseball$Year)]
nrow(subset(baseball,NumCompetitors==8))
table(baseball$NumCompetitors)
```

P3.1/3.2
```{r}
baseball$WorldSeries = as.numeric(baseball$RankPlayoffs == 1)
table(baseball$WorldSeries)
```


